Reduced growth due to belowground sink limitation is not fully explained by reduced photosynthesis

### 2016-12-12 manuscript/reviewer_response_tree_physiology.Rmd

---
output:
  word_document:
    reference_docx: manuscriptstyle_pce2.docx
  pdf_document: default
  html_document:
    number_sections: yes
---

```{r, results='asis', echo=FALSE, warning=FALSE}
library(knitr)
opts_knit$set(root.dir = '../')
```

Dr Ram Oren
\
Editor-in-Chief
\
Tree Physiology

\
Dear Professor Mott,

We are grateful for the opportunity to resubmit our manuscript. We thank the reviewers for providing insightful comments to improve our manuscript. We have responded in detail to each comment, which has improved the overall quality of the manuscript. Below, we have itemized our response to each comment from the Editor and both referees.

### **Referee #1**

\
*Reviewer: 1. pp35: "to utilize in growth and storage”*

\
*Response*: editorial change made

\
*Reviewer: 2. pp 40-45: Should be improved. Carbon losses are due to respiration and photorespiration.*

\
*Response*: **Change to net photosynthesis** instead of reviwer suggestion???

\
*Reviewer: 3. pp 63: “…the sink limitation of growth has been explored by manipulations of C source and sink activity…. Examples of …that manipulate C source activity include… and partial defoliation..” (partial defoliation should impose a sink limitation to growth?) Compensatory increases of An are commonly found, but not allowing a higher C net gain leading to sink limitation.*

\
*Response*:

\
*Reviewer: 4. pp 365: No changes in stomatal conductance during the experiment in the described field conditions (Figure 2)?*

\
*Response*: Soil moisture for each seedling were maitained at field capacity, throughout the duration of the experiment, in order to minimize affects of reduced water availability on photosynthesis and growth. The consistent lack of water stress on theses seedlings (see Discussion "Changes in growth and physiology under sink limitation") was an implement of the experimental design.

\
*Reviewer: 5. pp 373: Modelled cumulative net leaf gain could be displayed in a table? (e.g. Table 1)*

\
*Response*:

\
*Reviewer: 5. Table 1 and 2: Reorganize. Biomass, SLA, SRL and FRLD in one table, and chemistry parameters in a second table.*

\
*Response*

*Reviewer: 6. Figure 4: Significant differences between 5 and 15 L soil volumes? Are SE or SD displayed?*

\
*Response*: Bar plots of Figure 4 are means and standard errors which are described in figure caption. Yes, the 5 l containers had smaller rates of A~sat~ than either 15 or 20 L containers. The 5 l containers also had the lowest measured rates overall, indicative of a strong impact of a small container size on sink strength.  For simplicity, we report the large overall difference in Free seedlings and seedlings in containers in terms of rates of photosynthesis.

### **Referee #2**

\
*Reviewer: 1.  It is not clear whether trees used in this experiment come from a nursery or if they were lifted from the field*

\
*Response*: We have clarifed in "Experimental design" that seedlings were nursery seed grown in 40cm tall tubestock.

\
*Reviewer: 2. For Free saplings, it is not clear if roots could have expanded beyond the limits of the subplot or not.*

\
*Response*: We have clarified in "Experimental design" that roots were able to grow unrestricted below border material.

\
*Reviewer: 3 In relation to the “Photosynthetic parameters”, please, define Jmax and Vcmax.*

\
*Response*: We have defined both *V*~cmax~ and *J*~max~ in the section "Photosynthetic parameters"

\
*Reviewer: 4. In “Leaf, root and soil chemistry” indicate how plant samples were dried (temperature, drying time), how they were milled (type of mill used, model, etc.) and how soils were sieved (pore size) and milled (type of mill used, model, etc.)*

\
*Response*: We have clairified in “Leaf, root and soil chemistry” that samples were over dried at 60&degC to a constant mass and ball milled for analysis.We have also added the pore size, drying time and mill type to explanation of soil processing.

\
*Reviewer: 5.  Clairify how did soil volume treatments relate to the 61 digitalized seedlings, which include different *Eucalyptus* species? Were the 61 digitized seedlings grown in different soil volumes*

\
*Response*: The digitized seedlings used here are a collection of digitized plants from differing environments including understorey, glasshouse and plantation (see Duursma et al. 2012).  We selected 5 *Eucalpytus* species from this collection, which represent canopy structures of similar seedling types and sizes. Although the digitized plants are from seperate experiments, they allow us to estimate self shading of the *Eucalyptus* seedlings in this experiment.  As this experiment had an ~4 month duration, leaf counts and total leaf area of seedlings were generally small.  This resulted in only a modest adjsutment of the self-shading factor to total canopy carbon uptake.

\
*Reviewer: 6. Instead of using C concentration data (which they had), the authors assumed that C accounted for half of the final dry mass of seedlings. I think this is quite a strong assumption to make; differences in C concentration above 5% could be in place*

\
*Response*: **insert, I can implement the 48% from dryad wood C database, will change things very little but has citations**

\
*Reviewer: 7.  I would strongly recommend adopting a whole-plant approach and the consideration of all possible sources and sinks in mass-balance models. For example, root exudation can account for an important proportion of the total C budget of a plant, but yet this was not considered in the C budget calculations (nor even suggested as an important component to include in future models, see line 501)*

\
*Response*: The reviewer makes a very good point relating to how mass balance attempts should account for all possible fates of net C uptake. In this study, as in many, it is difficult to accurately measure every facet of C allocation throughout the plant and into the soil. It is entirely possible that aspects of belowground sink limitation could have impacted patterns of root exudation in these seedlings.  As we were not able many the possible fates of net C uptake, such as root exudation or woody tissue respiration, we utilized the mass balance approach to esimate the residual difference in C mass between the modelled seedling C uptake and harvested biomass. It is for this reason that we report differences in CUE, as a far more conservative approach in testing the coordiantion between photosynthesis and growth. For example, modelling the impacts of root exudation would require a priori knowledge of the effects of sink limitation/root restriction/container volume on this process. Consequently, we report the more conservative appraoch but give recommendations for future studies and models *here..here..here*

\
*Reviewer: 8.  I think future empirical and C models should evaluate the inter-relationship between the different sinks and sources AT THE WHOLE PLANT level, and not just the leaf level as indicated in lines 499-501.*

\
*Response*: We absolutely agree with the reviewer. We have made the editorial change to include the fate of C allocation at whole plant level, which includes the addition of root exudation from the previous comment.

\
*Reviewer: 9.  Lines 56-59. Yes, but beware the suggested approach also ignores part of the C budget of plants, because you are not accounting for root exudates.*

\
*Response*:

\
*Reviewer: 10.  Lines 71-72. Yes, indeed, C allocation belowground (both to roots and exudates) is enhanced under high CO2 supply.*

\
*Response*:  **add phillips/iversen sentence**??

\
*Reviewer: 11. Line 90. “Remove” or “removal”?*

\
*Response*: editorial change made

\
*Reviewer: 13.  Line 300. Use of prepositions. Replace “during” by “from”?*

\
*Response*: editorial change made

\
*Reviewer: 14.  Line 301. Remove “significantly” before “among”?*

\
*Response*: editorial change made

\
*Reviewer: 15.  Line 318. “…after variation in seedling biomass across treatments was factored in the analysis”, what do you mean by “factored”?*

\
*Response*: Replaced "factored" to "accounted for"

\
*Reviewer: 16.  Line 319. Remove “and” before “was”.*

\
*Response*: editorial change made

\
*Reviewer: 17.  Line 386. Or root exudates…*

\
*Response*:

\
*Reviewer: 18.  Line 461-462. Yes, and this could also explain the observed lower CUE in trees grown in small containers.*

\
*Response*:  **add exudation somewhere to CUE??**

\
*Reviewer: 19.  Line 513. Remove “to” before “predict”.*

\
*Response*: editorial change made

### 2016-12-12 manuscript/manuscript_pve.Rmd
started reviwer response

\
Alternatively, sink limitation can be investigated by manipulating plant tissue C sinks, for example by fruit remove or phloem girdling. In these studies, down regulation of *A*~n~ is often observed, and typically correlates with carbohydrate accumulation resulting from reduced tissue sink strength  [@iglesias2002regulation;@urban2007girdling; @haouari2013fruit]. However, reductions in *A*~n~ have also been attributed to biochemical limitations prior to carbohydrate accumulation [@nebauer2011photosynthesis], irreversible photo-oxidative damage [@duan2008photosynthetic] and stomatal limitation [@li2005photosynthesis]. 
Similar to C source manipulations, it is still unclear whether these impacts on *A*~n~ can fully explain the resulting impacts on growth.

@@

\
Forty-nine *Eucalyptus tereticornis* seedlings, 20 weeks old and approximately 40 cm tall in tube stock, were chosen from a single local Cumberland plain (western Sydney) cohort. Six additional seedlings were harvested before planting to measure initial leaf area and dry mass of leaves, stems and roots. Six container volumes were used ranging from 5 L to 35 L, with a 22.5 cm diameter, and lengths ranging from 15 to 100 cm. Containers were constructed of PVC pipe capped at the bottom end and were filled with local top soil. Soil in each container was packed to achieve a target soil bulk density that matched local soil conditions of 1.7 g cm^-3^. An Imidacloprid (BAYER CropScience) insecticide tablet was planted 5 cm below the roots of each seedling. Containers were planted flush with the soil surface inside metal sleeves, designed to minimize excess air space between the container and outside soil while also allowing for container removal. This allowed for soil temperatures in containers to reflect conditions of naturally planted (Free) seedlings. Each experimental block (n=7) contained a complete replicate set of six container volumes as well as one Free seedling, with 1 m^2^ spacing. For each Free seedling, used as the control, a 1 m^2^ subplot was excavated to the hard layer and replaced with the same soil used in each container. A border of root exclusion material was buried 0.25 m deep and extended 0.25 m above the ground surface around each subplot to exclude local vegetation, which was further kept out by periodic weeding.

@@

\
Photosynthetic CO~2~ response (AC~i~) curves were measured at 25 &deg;C on a random subset of each container size (n=3) after new leaves were first produced (March 13-14^th^, 2013) and prior to the final harvest (May 14-15^th^, 2013). Each AC~i~ curve was started at the reference [CO~2~] of 400 $\mu$l l^-1^ and then consisted of 12 additional steps from [CO~2~] of 50 to 1800 $\mu$l^-1^ at 25 &deg;C at saturating light. From these curves the photosynthetic parameters, *J*~max~ and *V*~cmax~, were quantified using the biochemical model of @farquhar1980biochemical and fit with the 'plantecophys' package [@duursma2015] in R [@RDevelopmentCoreTeam2011].

@@

##Leaf, root and soil chemistry
Each leaf used in gas exchange measurements and subsamples of harvested roots were dried to a constant mass and milled for analysis of nitrogen (N) content, $\delta$^13^C, and total non-structural carbohydrates (TNC). Pre-planting soil samples (n=6) and subsamples of soil from each container following harvest were sieved to remove organic material, air dried and milled for analysis of N. Nitrogen concentrations of leaf and soil samples were determined using a Carlo Erba CE1110 elemental analyzer and the percentage of N in samples was calculated by comparison with certified standards. Leaf $\delta$^13^C was analyzed  with an Delta V Advantage coupled to a Flash HT and Conflo IV isotope ratio mass spectrometer. Leaf samples were flash combusted at 1000&deg;C to convert to CO~2~, feed to the mass spectrometer and isotopic signatures are reported relative to the standard Vienna Pee Dee Belemnite scale.

@@

##Growth and morphology
Plant height, diameter and leaf area diverged between container volumes soon after start of the experiment (Figure `r allomfig`a-c). First to diverge was seedling leaf area, which significantly differed among soil volumes (*P* < 0.029) during the 5^th^ week of the experiment. Significant differences were then observed in both height (8^th^ week) and then diameter (9^th^ week) significantly among soil volumes (*P* < 0.045 & 0.035, respectively). The large reductions in height gain and total leaf area in smaller compared to larger containers continued throughout the experiment. In this field study, colder temperatures and reductions in total PPFD per day (Figure `r airvarsfig`) likely slowed the growth of seedlings in the final weeks of the experiment. Seedlings maintained diameter growth throughout the experiment, although growth was marginal in smaller soil volumes in the final month. Final seedling height significantly increased with increasing soil volume (*P* < 0.001). Increases in both final stem diameter and cumulative leaf area (both *P* < 0.001) were found with increasing soil volume. We did not find a continuous effect of container size; rather, plant size was similar among smaller containers, and differences were driven mainly by the largest container and the Free seedling treatments.

\
Total seedling biomass at harvest was significantly different across container volumes (*P* < 0.001) and between container treatments and Free seedlings (*P* < 0.001, Table `r pvetab1`). On average, harvested biomass of Free seedlings was 84 % higher than that of seedlings in containers. Plant biomass was positively correlated with total leaf area across all treatments (*R*^2^ = 0.97, *P* < 0.001). Differences in biomass partitioning to leaves, stems and roots were not different across soil volumes after variation in seedling biomass across treatments was factored in the analysis (Figure `r allocfig`a,b). The final harvest root:shoot biomass ratio and  was conserved across all treatments, with a slightly higher shoot than root mass ($\overline{x}$ = 0.904, 95 % CI = [0.846,1.119]). The ratio of leaf to fine root mass was also not different between treatments (Figure `r allocfig`c).

@@

\
Focusing only on empirical results in this study points to *A*~n~ driving reductions in seedling growth. Modelling cumulative net seedling C gain, however, suggests that other components of the C balance, beyond *A*~n~, are required to explain the observed seedling biomass response to sink manipulation. This finding is noteworthy, as partial accounting of the different mechanisms involved in plant C balance can lead to erroneous conclusions [@valentini2000respiration]. For example, the fraction of photosynthate used in respiration is known to vary depending on species and local environmental conditions [@Lambers2008], yet is often considered a static parameter in process based growth models. Thus, we agree with conclusions from Valentine and M&#228;kel&#228; [-@valentine2005bridging] that the problem with predicting tree growth is a problem in forecasting the assimilation and allocation of C and other constituents. Future empirical and modelling studies should focus on how feedbacks from sink activity affect both rates of *A*~n~ and the fate of C allocated to growth, respiration and C storage in leaves. However, the degree to which these mechanisms regulate the C available for growth will undoubtedly shift across different species and environmental conditions.

##Conclusions
With a unique field-based design we detected a massive effect of container volume on seedling growth, when compared with naturally planted seedlings. This finding is important as manipulations of plants grown in containers are often used to draw conclusions about growth and physiological principles, but how these results actually reflect field-grown plants is seldom studied. As seedlings maintained a conservative partitioning strategy, aboveground biomass was restricted in coordination with root restriction from limited soil volume. Photosynthesis was affected by both reduced nutrient uptake and the buildup of starch, both potential mechanism for downregulation of *A*~n~. Our combined empirical and modeling approach shows that when non-photosynthetic parameters were kept constant, changes in *A*~n~ were not able to fully to predict changes in growth, an important distinction often missed in studies that manipulate source/sink activity. It has been suggested that tissue sink activity strongly feedbacks onto source activity, causing growth to control *A*~n~ through the demand for C [@korner2013growth; @korner2015paradigm]. Our results imply that testing this hypothesis requires accurate modelling of C mass balance before addressing this debate.

>>>>

\
Alternatively, sink limitation can be investigated by manipulating plant tissue C sinks, for example by fruit removal or phloem girdling. In these studies, down regulation of *A*~n~ is often observed, and typically correlates with carbohydrate accumulation resulting from reduced tissue sink strength  [@iglesias2002regulation;@urban2007girdling; @haouari2013fruit]. However, reductions in *A*~n~ have also been attributed to biochemical limitations prior to carbohydrate accumulation [@nebauer2011photosynthesis], irreversible photo-oxidative damage [@duan2008photosynthetic] and stomatal limitation [@li2005photosynthesis]. 
Similar to C source manipulations, it is still unclear whether these impacts on *A*~n~ can fully explain the resulting impacts on growth.

@@

\
Forty-nine *Eucalyptus tereticornis* seedlings, 20 weeks old and approximately 40 cm tall in tube stock, were nursery seed grown from a single local Cumberland plain (western Sydney) cohort. Six additional seedlings were harvested before planting to measure initial leaf area and dry mass of leaves, stems and roots. Six container volumes were used ranging from 5 L to 35 L, with a 22.5 cm diameter, and lengths ranging from 15 to 100 cm. Containers were constructed of PVC pipe capped at the bottom end and were filled with local top soil. Soil in each container was packed to achieve a target soil bulk density that matched local soil conditions of 1.7 g cm^-3^. An Imidacloprid (BAYER CropScience) insecticide tablet was planted 5 cm below the roots of each seedling. Containers were planted flush with the soil surface inside metal sleeves, designed to minimize excess air space between the container and outside soil while also allowing for container removal. This allowed for soil temperatures in containers to reflect conditions of naturally planted (Free) seedlings. Each experimental block (n=7) contained a complete replicate set of six container volumes as well as one Free seedling, with 1 m^2^ spacing. For each Free seedling, used as the control, a 1 m^2^ subplot was excavated to the hard layer and replaced with the same soil used in each container. A border of root exclusion material was buried 0.25 m deep and extended 0.25 m above the ground surface around each subplot to exclude local vegetation, which was further kept out by periodic weeding. Roots of Free seedlings were allowed to grow unrestricted below the root exlusion border.

@@

\
Photosynthetic CO~2~ response (AC~i~) curves were measured at 25 &deg;C on a random subset of each container size (n=3) after new leaves were first produced (March 13-14^th^, 2013) and prior to the final harvest (May 14-15^th^, 2013). Each AC~i~ curve was started at the reference [CO~2~] of 400 $\mu$l l^-1^ and then consisted of 12 additional steps from [CO~2~] of 50 to 1800 $\mu$l^-1^ at 25 &deg;C at saturating light. From these curves, the maximum rate of Rubisco carboxylation (*V*~cmax~) and the maximum rate of electron transport for RuBP regeneration under saturating light (*J*~max~) were quantified using the biochemical model of @farquhar1980biochemical and fit with the 'plantecophys' package [@duursma2015] in R [@RDevelopmentCoreTeam2011].

@@

##Leaf, root and soil chemistry
Each leaf used in gas exchange measurements and subsamples of harvested roots were oven dried at 60&degC to a constant mass and ball milled for analysis of nitrogen (N) content, $\delta$^13^C, and total non-structural carbohydrates (TNC). Pre-planting soil samples (n=6) and subsamples of soil from each container following harvest were passed through a 1 mm sieve to remove organic material, air dried at 40&degC for 72 h and ball milled for analysis of N. Nitrogen concentrations of leaf and soil samples were determined using a Carlo Erba CE1110 elemental analyzer and the percentage of N in samples was calculated by comparison with certified standards. Leaf $\delta$^13^C was analyzed  with an Delta V Advantage coupled to a Flash HT and Conflo IV isotope ratio mass spectrometer. Leaf samples were flash combusted at 1000&deg;C to convert to CO~2~, feed to the mass spectrometer and isotopic signatures are reported relative to the standard Vienna Pee Dee Belemnite scale.

@@

##Growth and morphology
Plant height, diameter and leaf area diverged between container volumes soon after start of the experiment (Figure `r allomfig`a-c). First to diverge was seedling leaf area, which significantly differed among soil volumes (*P* < 0.029) from the 5^th^ week of the experiment. Significant differences were then observed in both height (8^th^ week) and then diameter (9^th^ week) among soil volumes (*P* < 0.045 & 0.035, respectively). The large reductions in height gain and total leaf area in smaller compared to larger containers continued throughout the experiment. In this field study, colder temperatures and reductions in total PPFD per day (Figure `r airvarsfig`) likely slowed the growth of seedlings in the final weeks of the experiment. Seedlings maintained diameter growth throughout the experiment, although growth was marginal in smaller soil volumes in the final month. Final seedling height significantly increased with increasing soil volume (*P* < 0.001). Increases in both final stem diameter and cumulative leaf area (both *P* < 0.001) were found with increasing soil volume. We did not find a continuous effect of container size; rather, plant size was similar among smaller containers, and differences were driven mainly by the largest container and the Free seedling treatments.

\
Total seedling biomass at harvest was significantly different across container volumes (*P* < 0.001) and between container treatments and Free seedlings (*P* < 0.001, Table `r pvetab1`). On average, harvested biomass of Free seedlings was 84 % higher than that of seedlings in containers. Plant biomass was positively correlated with total leaf area across all treatments (*R*^2^ = 0.97, *P* < 0.001). Differences in biomass partitioning to leaves, stems and roots were not different across soil volumes after variation in seedling biomass across treatments was accounted for in the analysis (Figure `r allocfig`a,b). The final harvest root:shoot biomass ratio was conserved across all treatments, with a slightly higher shoot than root mass ($\overline{x}$ = 0.904, 95 % CI = [0.846,1.119]). The ratio of leaf to fine root mass was also not different between treatments (Figure `r allocfig`c).

@@

\
Focusing only on empirical results in this study points to *A*~n~ driving reductions in seedling growth. Modelling cumulative net seedling C gain, however, suggests that other components of the C balance, beyond *A*~n~, are required to explain the observed seedling biomass response to sink manipulation. This finding is noteworthy, as partial accounting of the different mechanisms involved in plant C balance can lead to erroneous conclusions [@valentini2000respiration]. For example, the fraction of photosynthate used in respiration is known to vary depending on species and local environmental conditions [@Lambers2008], yet is often considered a static parameter in process based growth models. Thus, we agree with conclusions from Valentine and M&#228;kel&#228; [-@valentine2005bridging] that the problem with predicting tree growth is a problem in forecasting the assimilation and allocation of C and other constituents. Future empirical and modelling studies should focus on how feedbacks from sink activity affect both rates of *A*~n~ and the fate of C allocated at the whole plant level to growth, respiration, exudation and C storage. However, the degree to which these mechanisms regulate the C available for growth will undoubtedly shift across different species and environmental conditions.

##Conclusions
With a unique field-based design we detected a massive effect of container volume on seedling growth, when compared with naturally planted seedlings. This finding is important as manipulations of plants grown in containers are often used to draw conclusions about growth and physiological principles, but how these results actually reflect field-grown plants is seldom studied. As seedlings maintained a conservative partitioning strategy, aboveground biomass was restricted in coordination with root restriction from limited soil volume. Photosynthesis was affected by both reduced nutrient uptake and the buildup of starch, both potential mechanism for downregulation of *A*~n~. Our combined empirical and modeling approach shows that when non-photosynthetic parameters were kept constant, changes in *A*~n~ were not able to fully predict changes in growth, an important distinction often missed in studies that manipulate source/sink activity. It has been suggested that tissue sink activity strongly feedbacks onto source activity, causing growth to control *A*~n~ through the demand for C [@korner2013growth; @korner2015paradigm]. Our results imply that testing this hypothesis requires accurate modelling of C mass balance before addressing this debate.
